### 2022-04-01 ms/si.Rmd

---
title: "*Supplementary Material 1*"
subtitle: "Tables and Figures"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template_SI.docx
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Supplementary Material 1} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \setlength{\headheight}{13.6pt}
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

---
bibliography: 
- `r here::here("ms/si_references.yaml")`
- `r here::here("ms/si_references_other.yaml")`
---

```{r setup-si, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, results = 'hide')
# Load packages
library(ggupset)
library(flextable)
source(here::here("R/ms_packages.R"))
```

```{r load-functions-si, eval = params$doc_type == "manual"}
# Load functions (only if running interactively)
source("_targets.R")
```

```{r formatters-load-si}
# Load formatting functions defined in functions.R
knitr::read_chunk(
  here::here("R/functions.R"),
  labels = "formatters-si",
  from = "formatters start",
  to = "formatters end"
)
```

```{r run-formatters-si, eval = params$doc_type == "manual"}
source_chunk("formatters-si")
```

```{r formatters-si, eval = params$doc_type != "manual"}
```

```{r load-targets-si}
# Targets from FTOL workflow

# Specify path to FTOL cache
ftol_cache <- here::here("ftol_cache")

tar_load(
  c(
    acc_table_long,
    con_fossil_calibration_tips,
    con_fossil_calibrations_for_treepl,
    fossil_ferns_all,
    plastome_seqs_combined_filtered,
    plastome_tree,
    plastome_tree_rooted,
    ppgi_taxonomy,
    sanger_accessions_selection,
    sanger_sampling,
    sanger_con_tree_dated,
    sanger_con_tree_rooted,
    sanger_alignment_tbl,
    ts_fossil_calibration_tips
  ),
  store = ftol_cache
)

# Targets from MS workflow
tar_load(
  c(
    du_dates_all,
    family_stem_ages,
    fern_monophy_summ_tbl,
    fern_nonmono_gen,
    gb_species_by_year,
    other_dates,
    pteridocat_ppgi_diff_notes,
    pteridocat_ppgi_diff_notes_file,
    ref_files,
    spermato_trees,
    ts_family_stem_ages,
    accepted_species,
    ftol_ts_comp_mod
  ),
  store = here::here("_targets")
)
```

```{r load-captions, cache = FALSE}
# Source captions numbers from MS
# 
# So that caption numbers are in order of appearance in MS

# Run inline code when purling
options(knitr.purl.inline = TRUE)

# Make a temporary file to write out just inline R code from the SI Rmd
temp_file <- tempfile()

# Generate R script including inline R code from the ms Rmd
knitr::purl(here::here("ms/manuscript.Rmd"), output = temp_file)

# Trim this R script down to only functions that define figure and table captions
read_lines(temp_file) %>%
  magrittr::extract(
    str_detect(., "^figure\\(|^table\\(|^s_figure\\(|^s_table\\(|^figure_num\\(|^table_num\\(|^s_figure_num\\(|^s_table_num\\(")
  ) %>%
  unique %>%
  write_lines(temp_file)

# Source the code. This will load the caption functions in the right order.
source(temp_file, local = TRUE)

# Delete the temporary file
fs::file_delete(temp_file)
```

```{r load-pandoc, cache = FALSE}
# Specify bibliography and CSL files for pandoc
bib <- c(
  here::here("ms/si_references.yaml"),
  here::here("ms/si_references_other.yaml"),
  here::here("ms/si_fossils.bib"))
csl <- here::here("ms/frontiers.csl")
```

```{r extract-table-citations}
# Make list of references cited in pteridocat vs. ppgi differences table
# to include in bibliography
table_citations <- extract_citations(pteridocat_ppgi_diff_notes_file)$key %>%
  paste0("@", .) %>%
  paste0(collapse = ", ")
```

```{r plot-prep-load}
# Load the plot prep code from the main MS
knitr::read_chunk(
  here::here("ms/manuscript.Rmd"),
  labels = "plot-prep-main",
  from = "plot-prep start",
  to = "plot-prep end"
)
```

```{r run-plot-prep, eval = params$doc_type == "manual"}
source_chunk("plot-prep-main")
```

```{r plot-prep-main, eval = params$doc_type != "manual"}
```

---
nocite: |
  `r table_citations`
---

```{r genera-diff-make}
# FIXME: need to update ppgi csv file
# Add one missing nothogenus
ppgi_taxonomy_mod <-
  bind_rows(
    ppgi_taxonomy,
    tibble(
      genus = "Phlebosia",
      notes = "Nothogenus"
    )
  )

# Verify that all accepted genera are accounted for
accepted_species %>%
  count(genus, name = "n_species") %>%
  anti_join(select(ppgi_taxonomy_mod, genus, notes), by = "genus") %>%
  verify(nrow(.) == 0, success_fun = success_logical)

# Make table of pteridocat genera not in PPGI
pteridocat_not_in_ppgi <-
  accepted_species %>%
  count(genus, name = "n_species") %>%
  left_join(select(ppgi_taxonomy_mod, genus, notes), by = "genus") %>%
  assert(not_na, notes) %>%
  assert(is_uniq, genus) %>%
  filter(notes != "original PPGI 2016") %>%
  transmute(
    diff = "pteridocat not in PPGI",
    genus,
    n_species
  ) %>%
  # Add notes
  left_join(
    select(pteridocat_ppgi_diff_notes, -n_species),
    by = c("diff", "genus")
  ) %>%
  assert(not_na, note)

# Make table of PPGI genera not in pteridocat
ppgi_not_in_pteridocat <-
  ppgi_taxonomy_mod %>%
  filter(notes == "original PPGI 2016",
         class == "Polypodiopsida") %>%
  anti_join(accepted_species, by = "genus") %>%
  count(genus) %>%
  transmute(
    diff = "PPGI not in pteridocat",
    genus
  ) %>%
  # Add notes
  left_join(
    pteridocat_ppgi_diff_notes, by = c("diff", "genus")
  ) %>%
  assert(not_na, note)

genera_diff_table <- 
  bind_rows(pteridocat_not_in_ppgi, ppgi_not_in_pteridocat) %>%
  # Format references
  mutate(note = cite_pd(
    note, bib = bib, csl = csl, context = pteridocat_ppgi_diff_notes_file)) %>%
  # Check that all notes have references properly formatted
  # pandoc leaves a '?' if not
  assert(function(x) str_detect(x, "\\?", negate = TRUE), note)
```

```{r genera-diff-save, eval = params$doc_type == "doc"}
# Save if rendering to doc
write_csv(genera_diff_table, result_file(s_table("genera_diff"), "csv"))
```

**`r s_table("genera_diff")`**.
Differences in genera included as accepted in "pteridocat" taxonomic database v0.0.1 (https://github.com/fernphy/pteridocat) vs. PPGI [@PteridophytePhylogenyGroupI2016] (`r result_file(s_table("genera_diff"), "csv") %>% fs::path_file()`).
"diff" is type of difference (pteridocat genus not in PPGI or PPGI genus not in pteridocat).
"n_species" is the number of accepted species in each genus (pteridocat) or estimated number of species (PPGI).
References in "notes" column provided at end of Supplementary Materials.
"World Ferns" refers to @Hassler2022, which was modified to create pteridocat.

```{r fossils-make}
# Write out list of cited refs as "context" file
temp_context <- fs::path_abs(tempfile(fileext = ".txt"))

fossil_ferns_all %>%
  # Filter raw fossil data to just fossils used
  inner_join(
    unique(select(con_fossil_calibration_tips, n_fos)), by = "n_fos") %>%
  verify(nrow(.) == nrow(con_fossil_calibration_tips)) %>%
  select(where(is.character)) %>%
  pivot_longer(everything()) %>%
  filter(!is.na(value)) %>%
  pull(value) %>%
  write_lines(temp_context)

# Format references with pandoc
fossils_table <-
  fossil_ferns_all %>%
  # Filter raw fossil data to just fossils used
  inner_join(
    unique(select(con_fossil_calibration_tips, n_fos)), by = "n_fos") %>%
  verify(nrow(.) == nrow(con_fossil_calibration_tips)) %>%
  # Select columns
  select(
    n_fos, minimum_age, node_calibrated, fossil_taxon, 
    lower = lower_limit_of_oldest_stratigraphic_age,
    upper = upper_limit_of_oldest_stratigraphic_age,
    stratigraphic_age, 
    country, type_locality, formation,
    organs, specimen_holotype, collection,
    reference_raw = reference
  ) %>%
  # Format references
  mutate(reference = cite_pd(
    reference_raw, bib = bib, csl = csl, context = temp_context)) %>%
  select(-reference_raw) %>%
  # Make sure all references in the table have entries in the bibliography
  assert(function(x) str_detect(x, "\\?", negate = TRUE), reference)

fs::file_delete(temp_context)
```

```{r fossils-save, eval = params$doc_type == "doc"}
write_csv(fossils_table, result_file(s_table("fossils"), "csv"))
```

**`r s_table("fossils")`**.
Fossils used as calibration points for molecular dating (`r result_file(s_table("fossils"), "csv") %>% fs::path_file()`).
"n_fos" is the fossil ID number used in the "ferncal" database v1.0.0 (https://github.com/fernphy/ferncal).
"minimum_age" is the age used for calibration (Ma); "lower" is the lower limit of the oldest stratigraphic age for the fossil (Ma); "upper" is the upper limit of the oldest stratigraphic age for the fossil (Ma).
For detailed notes and full references for each fossil, see Appendix S2.

```{r accs-save, eval = params$doc_type == "doc"}
write_csv(acc_table_long, result_file(s_table("accs"), "csv"))
```

**`r s_table("accs")`**.
GenBank accessions used in this study (`r result_file(s_table("accs"), "csv") %>% fs::path_file()`).
"species" is the species name as it appears in the phylogenetic tree.
"seq_len" is the total number of bases excluding missing bases ("?", "N", or "-") and only includes bases belonging to loci used in this study (not the entire accession).
"sci_name" is the scientific name in the pteridocat taxonomic database.
"ncbi_name" is the scientific name used in the NCBI taxonomic database version 2022-02-01.
"ncbi_taxid" is the NCBI taxonomy database unique identifier.
"outgroup_status" is "in" for ingroup taxa (ferns) and "out" for outgroup taxa.

\newpage

```{r join-by, results = "markup"}
# Make table showing method of joining accessions across loci
sanger_accessions_selection %>% 
  # Remove outgroup and plastome species
  anti_join(
    unique(select(plastome_seqs_combined_filtered, species)),
    by = "species") %>%
  assert(not_na, join_by, species) %>%
  assert(is_uniq, species) %>%
  tabyl(join_by) %>%
  arrange(n) %>%
  adorn_totals() %>%
  mutate(
    join_by = str_to_sentence(join_by) %>%
      str_replace_all("Monophy", "Monophyletic"),
    n = number(n),
    percent = case_when(
      join_by != "Total" ~ percent(percent),
      TRUE ~ ""
    )
  ) %>%
  rename(
    "Method" = join_by,
    `%` = percent) %>%
  flextable() %>%
  italic(1,2, part = "header") %>%
  autofit()
```

**`r s_table("join_by")`**.
Methods used to join accessions across loci for Sanger dataset.
Count is number of species.

\newpage

```{r monophy, results = "markup"}
fern_monophy_summ_tbl %>%
  mutate(
    tax_level = str_to_sentence(tax_level),
    percent = percent(percent, accuracy = 0.1) %>%
      str_replace_all("100\\.0", "100")) %>%
  select(
    `Taxonomic level` = tax_level, Monophyly = monophyly,
    n, `%` = percent) %>%
  flextable() %>%
  italic(1,3, part = "header") %>%
  autofit()
```

**`r s_table("monophy")`**.
Monophyly by taxonomic level for ferns in FTOL.

\newpage

```{r nonmono-gen, eval = params$doc_type == "doc"}
fern_nonmono_gen %>%
  arrange(family, subfamily, genus) %>%
  write_csv(result_file(s_table("nonmono_gen"), "csv"))
```

**`r s_table("nonmono_gen")`**.
Non-monophyletic fern genera in FTOL (`r result_file(s_table("nonmono_gen"), "csv") %>% fs::path_file()`).
"number_tips" is the number of tips (species) in the genus.
"delta_tips" is the number of tips that do not belong to the genus but are included in the clade descending from the most common recent ancestor of the genus.
"number_intruders" is the number of intruding taxa into the genus.
"number_outliers" is the number of tips (species) belonging to the genus that appear in other clades.
Status of intruders and outliers was determined using the default cutoff level (0.5) in the R package "MonoPhy" [@Schwery2016].

\newpage

```{r fossil-map-make}
# Make tibble with node number and label for ggtree
node_labs_tbl <-
  con_fossil_calibrations_for_treepl %>%
  transmute(
    fossil_label = str_remove_all(min, "min = "),
    node = map2_dbl(taxon_1, taxon_2, ~ape::getMRCA(sanger_con_tree_rooted, c(.x, .y))),
    calibration_type = case_when(
      !is.na(min) & is.na(max) ~ "minimum",
      !is.na(min) & !is.na(max) ~ "fixed",
      TRUE ~ NA_character_
    )
  ) %>%
  assert(not_na, calibration_type)

# Make plot
fossil_map_plot <- 
  ggtree(
    sanger_con_tree_rooted,
    branch.length = "none", size = 0.1, color = "grey30") %<+% node_labs_tbl +
  # Add node points
  geom_nodepoint(
    aes(color = calibration_type), size = 0.6) +
  scale_color_manual(
    "Calibration type",
    values = c(
      minimum = palette_okabe_ito("vermillion"), 
      fixed = palette_okabe_ito("skyblue")),
    na.value = "transparent"
  ) +
  # Add node labels
  geom_text(aes(label = fossil_label), hjust = 0, size = 1.0, color = "grey20") +
  # Add tip labels
  geom_tiplab(aes(label = label), size = 0.5, color = "grey30") +
  theme(legend.position = "bottom")
```

```{r fossil-map-save, eval = params$doc_type == "doc"}
ggsave(
  plot = fossil_map_plot,
  file = (result_file(s_figure("fossil_map"), "pdf")),
  height = 60, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("fossil_map")`**.
Consensus phylogeny (Sanger dataset) with node labels indicating placement of fossil constraints (`r result_file(s_figure("fossil_map"), "pdf") %>% fs::path_file()`).
Number for each constraint is age in Ma.

\newpage

```{r locus-upset, fig.width = 7, fig.height = 5, eval = params$doc_type != "manual"}
# Make tibble with one row per locus per species
sanger_alignment_tbl %>%
  mutate(taxon = map(align_trimmed, rownames)) %>%
  select(locus = target, species = taxon) %>%
  unnest(cols = species) %>%
  # Exclude outgroups
  left_join(
    unique(select(sanger_sampling, species, outgroup)), by = "species") %>%
  filter(outgroup == FALSE) %>%
  select(-outgroup) %>%
  group_by(species) %>%
  summarize(locus = list(locus)) %>%
  ggplot(aes(x = locus)) +
  geom_bar() +
  labs(x = "Locus combination", y = "Count (num. species)") +
  scale_x_upset(n_intersections = 20)
```

**`r s_figure("locus-upset")`**.
Frequency of locus combinations by species, Sanger dataset.
Only top 20 combinations shown.

\newpage

```{r backbone-full-plot}
phy <- plastome_tree_rooted

# Make tibble for printing BS values
bs_tibble <- make_bs_tbl(phy) %>%
  # only print values < 100
  mutate(bs = case_when(
    bs == 100 ~ "",
    TRUE ~ as.character(bs)
  ))

# Make plot
bb_full_plot <- 
  ggtree(phy, size = 0.2, color = "grey10") %<+% 
  bs_tibble +
  # now we can show BS values using geom_text()
  geom_text(
    aes(label = bs), hjust = -.25, size = 1.0,
    color = "grey50") +
  # add tip labels
  geom_tiplab(aes(label = label), size = 1.0)
```

```{r backbone-full-save, eval = params$doc_type == "doc"}
ggsave(
  plot = bb_full_plot,
  file = (result_file(s_figure("backbone_full"), "pdf")),
  height = 20, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("backbone_full")`**.
Consensus plastome (backbone) phylogeny (`r result_file(s_figure("backbone_full"), "pdf") %>% fs::path_file()`).
Values at nodes are bootstrap support (%); only values less than 100% are shown.

\newpage

```{r bs-plot}
phy <- sanger_con_tree_rooted

# Make tibble for printing BS values
bs_tibble <- make_bs_tbl(phy) %>%
  mutate(bs = bs*0.01) # so percentages print correctly

# Make plot
bs_plot <- 
  ggtree(
    phy,
    size = 0.1, color = "grey30") %<+% bs_tibble +
  # Add node points
  geom_nodepoint(
    aes(color = bs), size = 0.1) +
  scale_color_viridis_c(
    "BS",
    option = "B", na.value = "transparent",
    labels = scales::percent
    ) +
  # Add tip labels
  geom_tiplab(aes(label = label), size = 0.5, color = "grey30") +
  theme(legend.position = "bottom")
```

```{r bs-save, eval = params$doc_type == "doc"}
ggsave(
  plot = bs_plot,
  file = result_file(s_figure("bs"), "pdf"),
  height = 60, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("bs")`**.
Consensus phylogeny (Sanger dataset) with node labels showing bootstrap support (%) (`r result_file(s_figure("bs"), "pdf") %>% fs::path_file()`).
Darker colors indicate lower support.
The backbone (plastome) phylogeny was used as a constraint tree, so all clades present in the backbone phylogeny tree receive 100% bootstrap support.

\newpage

```{r du-ages, fig.width = 7, fig.height = 5, eval = params$doc_type != "manual"}
# Helper function for subsetting dates to get into
# almost-long format
subset_ages <- function(other_dates, author_sel) {
  other_dates %>%
    select(family, contains(glue("{author_sel}_"))) %>%
    rename_with(~str_remove_all(., glue("{author_sel}_"))) %>%
    mutate(author = author_sel)
}

# Filter Du ages to just stem families
du_dates_stem_fam <-
  du_dates_all %>%
  filter(
    str_detect(affinities, "eae$"),
    affinities_group == "stem",
    str_detect(author, "^r")) %>%
  select(family = affinities, author, median, low, high)

# Make tibble of Du
du_schemes <- du_dates_stem_fam %>%
  select(scheme = author) %>%
  unique() %>%
  mutate(
    code = str_split(scheme, "_") %>%
      map_chr(last) %>%
      str_to_upper()) %>%
  arrange(code)

# Combine dates in almost-long format:
# one row per family/author combination,
# cols for median, high, and low date
all_ages_with_du <-
  bind_rows(
    du_dates_stem_fam,
    subset_ages(other_dates, "ts"),
    subset_ages(other_dates, "rothfels"),
    select(other_dates, family, median = schuettpelz_best) %>%
      mutate(author = "schuettpelz"),
    transmute(
      family_stem_ages,
      family,
      author = "current",
      median = age
    )
  ) %>%
  filter(family %in% du_dates_stem_fam$family) %>%
  left_join(
    unique(select(sanger_sampling, family, major_clade)), by = "family") %>%
  left_join(clade_tibble, by = "major_clade") %>%
  # (some median dates are NA if that family wasn't included in a author)
  assert(not_na, family, major_clade, code, author) %>%
  # Set family as a factor ordered by current age estimate
  left_join(family_stem_ages, by = "family") %>%
  mutate(family = fct_reorder(family, age)) %>%
  select(-age)

# Define shapes and colors for points
du_ages_colors <- c(
  current = palette_okabe_ito(order = "vermillion"),
  rothfels = palette_okabe_ito(order = "bluishgreen"),
  schuettpelz = palette_okabe_ito(order = "blue"),
  ts = palette_okabe_ito(order = "orange"),
  rep(
    palette_okabe_ito(order = "gray"),
    length(du_schemes$scheme)) %>%
    set_names(du_schemes$scheme)
)

du_ages_shapes <- c(
  current = 16, # circle
  rothfels = 16,
  schuettpelz = 16,
  ts = 16,
  seq_along(du_schemes$scheme) %>%
    subtract(1) %>%
    set_names(du_schemes$scheme)
)

# Cross doesn't work because we can't distinguish error bars
du_ages_shapes[du_ages_shapes == 3] <- 6

# Format names of papers for printing
du_ages_sampling <-
  tribble(
    ~source, ~year, ~label,
    "schuettpelz", 2009, "Schuettpelz and Pryer 2009",
    "rothfels", 2015, "Rothfels et al. 2015",
    "ts", 2016, "Testo and Sundue 2016",
    "current", 2021, "This study",
  ) %>%
  bind_rows(
    du_schemes %>%
      mutate(
        year = 2020 + 0.1*seq_along(du_schemes$scheme),
        label = paste("Du et al. 2021 scheme", code)
      ) %>%
      transmute(source = scheme, year, label)
  ) %>%
  mutate(source = fct_reorder(source, year))

# Format y-axis text: show major clade in parentheses
family_labs_tbl <-
  all_ages_with_du %>%
  select(family, code) %>%
  unique() %>%
  mutate(label = glue("{family} ({code})"))

family_labs <- family_labs_tbl$label %>%
  set_names(., family_labs_tbl$family)

ggplot(all_ages_with_du) +
  geom_pointrange(
    aes(
      x = median, xmin = low, xmax = high,
      y = family,
      color = author, shape = author),
    position = position_dodge(width = 0.5)) +
  scale_shape_manual(
    values = du_ages_shapes,
    limits = du_ages_sampling$source, 
    labels = du_ages_sampling$label) +
  scale_color_manual(
    values = du_ages_colors,
    limits = du_ages_sampling$source, 
    labels = du_ages_sampling$label) +
  scale_y_discrete(labels = family_labs, expand = expansion(add = 0.7)) +
  scale_x_reverse("Age (millions of years)", expand = c(0,0)) +
  coord_geo(
    dat = list("periods", "eras"),
    pos = list("b", "b"),
    abbrv = list(TRUE, FALSE),
    xlim = c(250, 0),
    fill = "transparent",
    color = "grey20",
    lab_color = "grey30",
    height = unit(1, "lines"),
    expand = TRUE,
    size = 3) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    legend.title = element_blank()
  )
```

**`r s_figure("du-ages")`**.
Stem age of fern families (Ma) estimated by selected studies, only including families that were sampled across all studies.
For studies that used methods with confidence intervals, error bars indicate lower and upper 95% highest posterior density levels and point indicates median [@Rothfels2015a; @Testo2016a; @Du2021].
For other studies, point indicates best (most likely) estimate [@Schuettpelz2009\; this study].
@Du2021 used six different dating schemes; for details, see that paper.
Codes in parentheses after family names indicate major clade as in `r figure("ftol")`.
Period name abbreviations as in `r figure("divtimes")`.

\newpage

```{r comp-fossils, fig.width = 7, fig.height = 9, eval = params$doc_type != "manual"}
ts_fossil_calibration_tips %>%
  select(minimum_age, node_calibrated) %>%
  mutate(dataset = "ts") %>%
  bind_rows(
    con_fossil_calibration_tips %>%
      select(minimum_age, node_calibrated) %>%
      mutate(dataset = "current")
  ) %>%
  mutate(node_calibrated = fct_reorder(node_calibrated, minimum_age)) %>%
  ggplot(aes(x = minimum_age, y = node_calibrated, fill = dataset)) +
  geom_col(position = position_dodge()) +
  scale_x_continuous("Age (millions of years)", expand = expansion(mult = c(0,0.1))) +
  theme_bw() +
  scale_fill_manual(
    values = c(current = "grey30", ts = "grey80"),
    labels = c(current = "This Study", ts = "Testo and Sundue 2016")
  ) +
  theme(
    axis.title.y = element_blank(),
    legend.title = element_blank(),
    axis.text.y = element_text(size = 6),
    legend.position = "bottom"
  )
```

**`r s_figure("comp-fossils")`**.
Age of nodes constrained by fossils in this study vs. @Testo2016a. Wide bars indicate the node was only constrained in one study; narrow bars indicate the the node was constrained in both studies.

\newpage

```{r comp-stem-ages, fig.width = 7, fig.height = 9, eval = params$doc_type != "manual"}
# Combine family stem ages into almost-long format:
# one row per family/study combination,
# cols for median, high, and low date
all_family_stem_ages <-
  bind_rows(
    subset_ages(other_dates, "ts"),
    transmute(
      ts_family_stem_ages,
      family,
      author = "current",
      median = age
    )
  ) %>%
  left_join(
    unique(select(sanger_sampling, family, major_clade)), by = "family") %>%
  left_join(clade_tibble, by = "major_clade") %>%
  # (some median dates are NA if that family wasn't included in a study)
  assert(not_na, family, major_clade, code, author) %>%
  # Set family as a factor ordered by current age estimate
  left_join(ts_family_stem_ages, by = "family") %>%
  mutate(family = fct_reorder(family, age)) %>%
  select(-age)

# Format y-axis text: show major clade in parentheses
family_labs_tbl <-
  all_family_stem_ages %>%
  select(family, code) %>%
  unique() %>%
  mutate(label = glue("{family} ({code})"))

family_labs <- family_labs_tbl$label %>%
  set_names(., family_labs_tbl$family)

# Make plot
ggplot(all_family_stem_ages) +
  geom_pointrange(
    aes(
      x = median, xmin = low, xmax = high,
      y = family,
      color = author, shape = author),
    position = position_dodge(width = 0.5)) +
  scale_shape_manual(
    values = ages_shapes[c("ts", "current")],
    limits = c("ts", "current"), 
    labels = ages_sampling %>% 
      filter(source %in% c("current", "ts")) %>%
      arrange(source) %>%
      pull(label)) +
  scale_color_manual(
    values = ages_colors[c("ts", "current")],
    limits = c("ts", "current"), 
    labels = ages_sampling %>% 
      filter(source %in% c("current", "ts")) %>%
      arrange(source) %>%
      pull(label)) +
  scale_y_discrete(labels = family_labs, expand = expansion(add = 0.7)) +
  scale_x_reverse("Age (millions of years)", expand = c(0,0)) +
  coord_geo(
    dat = list("periods", "eras"),
    pos = list("b", "b"),
    abbrv = list(TRUE, FALSE),
    xlim = c(500, 0),
    fill = "transparent",
    color = "grey20",
    lab_color = "grey30",
    height = unit(1, "lines"),
    expand = TRUE,
    size = 3) +
  guides(shape = guide_legend(nrow = 2)) +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "bottom"
  )
```

**`r s_figure("comp-stem-ages")`**.
Comparison of stem age of fern families (Ma) estimated by @Testo2016a and those estimated using methodology (treePL) and phylogeny (FTOL) of the current study with the fossil constraints of @Testo2016a.
For @Testo2016a, error bars indicate lower and upper 95% highest posterior density levels and point indicates median
For this study, point indicates best (most likely) estimate.
Codes in parentheses after family names indicate major clade as in `r figure("ftol")`.
Period name abbreviations as in `r figure("divtimes")`.

\newpage

```{r comp-stem-points, fig.width = 6, fig.height = 5, eval = params$doc_type != "manual"}
# For checking that model data is same as plot data
mod_fitted_data <-
ftol_ts_comp_mod %>%
  broom::augment() %>%
  arrange(ftol)

mod_fitted_data_ts <-
ftol_ts_comp_mod %>%
  broom::augment() %>%
  arrange(ts)

# Make dataset for plotting
comp_plot_data <-
all_family_stem_ages %>%
  filter(author == "ts") %>%
  select(family, ts = median, ts_high = high, ts_low = low) %>%
  inner_join(
    filter(all_family_stem_ages, author == "current") %>%
      select(family, ftol = median),
    by = "family"
  ) %>%
  # Double check that model data is same as plot data
  arrange(ftol) %>%
  verify(all(.$ftol == mod_fitted_data$ftol)) %>%
  arrange(ts) %>%
  verify(all(.$ts == mod_fitted_data_ts$ts))

# Extract R2 and P-value for linear mod
stats <- ftol_ts_comp_mod %>%
  broom::glance() %>%
  mutate(label = glue::glue(
    "*R*<sup>2</sup> = {number(r.squared, accuracy = 0.01)}\n
    {pval} = {sci_number(p.value)}"
  )) %>%
  pull(label)

# Extract fitted line for linear mod
fitted <-
broom::tidy(ftol_ts_comp_mod) %>%
  mutate(term = janitor::make_clean_names(term)) %>%
  split(.$term)

ggplot(comp_plot_data, aes(y = ts, x = ftol)) +
  geom_errorbar(aes(ymax = ts_high, ymin = ts_low)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "grey60", lty = 2) +
  geom_abline(
    intercept = fitted$intercept$estimate,
    slope = fitted$ts$estimate, color = "grey20"
  ) +
  theme_bw() +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "This study", y = "Testo and Sundue (2016)") +
  annotate(
    "richtext", x = 0, y = Inf, 
    label.color = NA,
    label = stats, hjust = 0, vjust = 1)
```

**`r s_figure("comp-stem-points")`**.
Comparison of stem age of fern families estimated by @Testo2016a and those estimated using methodology (treePL) and phylogeny (FTOL) of the current study with the fossil constraints of @Testo2016a.
All units in millions of years (Ma).
For @Testo2016a, error bars indicate lower and upper 95% highest posterior density levels and point indicates median.
For this study, point indicates best (most likely) estimate.
Dashed indicates 1:1 relationship.
Solid line indicates linear model fit.

\newpage

```{r ltt, fig.width = 7, fig.height = 6, eval = params$doc_type != "manual"}
# Make lineages through time (LTT) plot comparing Angiosperms and polypods

## Angiosperms ##
# FIXME: if we are really going to use this, move it to _targets.R
# Li 2019 data
li_species_tree <- ape::read.tree(here::here("_targets/user/data_raw/2881_dating_Species_extracted.tree"))
li_og <- read_lines(here::here("_targets/user/data_raw/2881_dating_Species_extracted_og.txt"))
li_angio_tree <- ape::drop.tip(li_species_tree, li_og)

li_tt_data <- make_ltt_tib(li_angio_tree, "Li_2019")

# Define outgroup taxa
og <- c(
  "Pinales_Pinaceae_Pinus_spp", "Gnetales_Welwitschiaceae_Welwitschia_spp",
  "Cycadales_Cycadaceae_Cycas_spp", "Gnetales_Gnetaceae_Gnetum_spp", 
  "Pinales_Cupressaceae_Metasequoia_glyptostroboides", 
  "Ginkgoales_Ginkgoaceae_Ginkgo_spp")

# Drop outgroups
angio_trees <-
  map(spermato_trees, ~ape::drop.tip(., og))

# Get data for LTT plot
angio_tt_data <-
  map2_df(angio_trees, names(angio_trees), ~make_ltt_tib(.x, .y))

## Polypods ##

# Get data for LTT plot
poly_tt_data <-
  sanger_sampling %>%
  filter(order == "Polypodiales") %>%
  pull(species) %>% 
  ape::keep.tip(sanger_con_tree_dated, .) %>%
  make_ltt_tib(., "Polypodiales")

## Combine and plot ##
tt_data <-
  bind_rows(angio_tt_data, poly_tt_data, li_tt_data)

ltt_dataset_cols <- c(
  CC_complete = palette_okabe_ito("orange"),
  CC_conservative = palette_okabe_ito("skyblue"),
  RC_complete = palette_okabe_ito("bluishgreen"),
  RC_conservative = palette_okabe_ito("blue"),
  UC_complete = palette_okabe_ito("vermillion"),
  UC_conservative = palette_okabe_ito("reddishpurple"),
  Li_2019 = palette_okabe_ito("yellow"),
  Polypodiales = palette_okabe_ito("black")
)

raw_count <- ggplot(tt_data, aes(x = time, y = N, color = dataset)) +
  geom_step() +
  scale_y_continuous("Number of lineages") +
  scale_x_reverse("Age (millions of years)") +
  coord_geo(
    dat = list("periods", "eras"),
    pos = list("b", "b"),
    abbrv = list(TRUE, FALSE),
    xlim = c(310, 0),
    fill = "transparent",
    color = "grey20",
    lab_color = "grey30",
    height = unit(1, "lines"),
    expand = FALSE,
    size = 3) +
  scale_color_manual("Dataset", values = ltt_dataset_cols) +
  theme_bw()

ggplot(tt_data, aes(x = time, y = log_n_percent, color = dataset)) +
  geom_step() +
  scale_y_continuous("Log number of lineages (as percent)", labels = scales::percent) +
  scale_x_reverse("Age (millions of years)") +
  coord_geo(
    dat = list("periods", "eras"),
    pos = list("b", "b"),
    abbrv = list(TRUE, FALSE),
    xlim = c(310, 0),
    fill = "transparent",
    color = "grey20",
    lab_color = "grey30",
    height = unit(1, "lines"),
    expand = FALSE,
    size = 3) +
  scale_color_manual("Dataset", values = ltt_dataset_cols) +
  theme_bw() +
  theme(legend.position = "bottom")
```

**`r s_figure("ltt")`**.
Lineage through time (LTT) plots for the fern order Polypodiales (black; this study) and angiosperms [colored lines\; @Li2019; @Ramirez-Barahona2020].
Angiosperm datasets labeled with "CC", "RC", or "UC" correspond to the six analysis schemes of @Ramirez-Barahona2020; for details, see that paper.
Log number of lineages shown as percent of total to enable comparison across trees.

\newpage

## References